---
title: "cancer_data_plots"
author: "Arnab Sarkar(as3uj)"
date: "23/11/2018"
output: html_document
---

```{r}
library(tidyverse)
load("D:/Courses/STAT 6021 Fall 2018/Project/Repository/cancer-prediction-project/train_v2.RData")
load("D:/Courses/STAT 6021 Fall 2018/Project/Repository/cancer-prediction-project/test_v2.RData")
attach(train)
```

Getting summary of all columns for the cancer train data-
```{r}
summary(train)
```


```{r}
pairs(~TARGET_deathRate+medIncome+povertyPercent,data=train)
```
target death rate appears to fall in a curvilinear manner with increase in median income (include:Yes?)
target death rate increases(slightly) linearly with povertyPercent (include:Yes?)

```{r}
plot(popEst2015,TARGET_deathRate)
```
There does not seem to be conclusive relationship between Population of County and death rate.(include:No)

```{r}
plot(studyPerCap,TARGET_deathRate)
```
No conclusive pattern between death rate and Per capita number of cancer-related clinical trials.(include:No)

```{r}
par(mar=c(8,5,1,1))
boxplot(TARGET_deathRate~binnedInc,data=train, main="Median Income bracket boxplots",
    ylab="death rate",las=2)
```
Death rate decreases with increase in median Income bracket. (include:Yes)

```{r}
pairs(~TARGET_deathRate+MedianAge+MedianAgeMale+MedianAgeFemale,data=train)
```

There appear to be lot incorrect/extreme values for MedianAge(>300).Correcting these would generate
a better plot.
Death rate does not have any conclusive relationship with MedianAgeMale. (include:No?) 
Death rate does not have any conclusive relationship with MedianAgeFeMale. (include:No?)

```{r}
plot(MedianAge,TARGET_deathRate,xlim=c(0,100))
```
death rate does not seem to have any relationship with MedianAge(after limiting x to 100) (include:No?)

```{r}
pairs(~TARGET_deathRate+AvgHouseholdSize+PercentMarried,data=train)
```

There does not seem to be apparent relationship between AvgHouseholdSize and death rate.Also some obs have AvgHouseholdSize set as 0. (include:No)
There appears to be a very slight decrease in death rate with PercentMarried. (include:Yes?)

```{r}
pairs(~TARGET_deathRate+PctNoHS18_24+PctHS18_24+PctSomeCol18_24+PctBachDeg18_24,data=train)
```

None of the four 18-24 age group variables show any distinct pattern w.r.t death rate.

Let's try converting these 4 into a single categorical variable depending on majority percentage-
```{r}
train$maj18_24<-NA
```


```{r}
train[which(train$PctNoHS18_24>=train$PctHS18_24&train$PctNoHS18_24>=train$PctSomeCol18_24&train$PctNoHS18_24>=train$PctBachDeg18_24),'maj18_24']<-'PctNoHS18_24'
train[which(train$PctHS18_24>=train$PctNoHS18_24&train$PctHS18_24>=train$PctSomeCol18_24&train$PctHS18_24>=train$PctBachDeg18_24),'maj18_24']<-'PctHS18_24'
train[which(train$PctSomeCol18_24>=train$PctNoHS18_24&train$PctSomeCol18_24>=train$PctHS18_24&train$PctSomeCol18_24>=train$PctBachDeg18_24),'maj18_24']<-'PctSomeCol18_24'
train[which(train$PctBachDeg18_24>=train$PctNoHS18_24&train$PctBachDeg18_24>=train$PctHS18_24&train$PctBachDeg18_24>=train$PctSomeCol18_24),'maj18_24']<-'PctBachDeg18_24'
```

Converting to factor and plotting as boxplots
```{r}
par(mar=c(9,5,1,1))
train$maj18_24<-as.factor(train$maj18_24)
boxplot(TARGET_deathRate~maj18_24,data=train, main="maj18_24 boxplots",
    ylab="death rate",las=2)
```

There does not seem to much variation between the categories of maj18_24.

```{r}
pairs(~TARGET_deathRate+PctHS25_Over+PctBachDeg25_Over,data=train)
```

There seems to be slight increase in death rate with higher percentage of 25 and over HS graduate(include:yes?)
There is a curvilinear relationship(decreasing) between deathrate and percentage over 25 with bachelor degree.(include:yes)

```{r}
pairs(~TARGET_deathRate+PctEmployed16_Over+PctUnemployed16_Over,data=train)
```

death rate appears to have linear relationship with both 16 over variables.
It decreases slightly with increase in PctEmployed16_Over (Include:Yes?)
Increases slightly with increase in PctUnemployed16_Over (Include:Yes).

```{r}
pairs(~TARGET_deathRate+PctPrivateCoverage+PctPrivateCoverageAlone+PctPublicCoverage+PctPublicCoverageAlone,data=train)
```

There appears to be linear relationship with private coverage variables and deathrate as well as between
public coverage and deathrate. (Include all: Yes?)
Not sure if above 4 variables can be combined in a meaningful manner(as they may have overlap)

The private employer coverage(PctEmpPrivCoverage.greater.than.mean) can be represented as boxplots
```{r}
par(mar=c(9,5,1,1))
boxplot(TARGET_deathRate~PctEmpPrivCoverage.greater.than.mean,data=train, main="employer private coverage boxplots",
    ylab="death rate",las=2)
```

People without employer private coverage seem to have a higher death rate, though spread seems to be same. (include: Yes?)

```{r}
pairs(~TARGET_deathRate+PctWhite+PctBlack+PctAsian+PctOtherRace,data=train)
```

There seems to no trend visible between death rate and race (Include:No?)

We can try combining the 4 variables into a single variable for majority race in county,
however asians and other race percentages are very small and dont cross even 40%.We can
combine the percentages of asians with other races-

```{r}
train$maj_race<-NA
train[which(train$PctWhite>=train$PctBlack&train$PctWhite>=(train$PctAsian+train$PctOtherRace)),'maj_race']<-'White'
train[which(train$PctBlack>=train$PctWhite&train$PctBlack>=(train$PctAsian+train$PctOtherRace)),'maj_race']<-'Black'
train[which((train$PctAsian+train$PctOtherRace)>=train$PctWhite&(train$PctAsian+train$PctOtherRace)>=train$PctBlack),'maj_race']<-'Other'

train$maj_race<-as.factor(train$maj_race)

par(mar=c(9,5,1,1))
boxplot(TARGET_deathRate~maj_race,data=train, main="Race demographic boxplots",
    ylab="death rate",las=2)
```

Above box plot shows that counties with majority black population have higher cancer death rate,
compared to white or others(least death rate due to cancer)
However be wary of the fact that most counties have majority white population and for them there
is still a lot of variation(though middle 50% of white counties appear to have death rate somewhere
in between the median black and median other death rates)


```{r}
plot(PctMarriedHouseholds,TARGET_deathRate)
plot(PercentMarried,TARGET_deathRate)
plot(PercentMarried,PctMarriedHouseholds)
```
Plotting PctMarriedHouseholds against deathrate shows marginal decrease with increase in PctMarriedHousehold.(Include:Yes?)
Comparing with PercentMarried variable which was explored earlier,both have similar relationship with deathrate and we may find both these variables to be co-linear(see plot 3 above) as well(may have to drop one)

```{r}
plot(BirthRate,TARGET_deathRate)
```
Birthrate of county shows no apparent relationship to cancer death rate. (Include: No)
